Project

### 2018-11-22 project.Rmd
change

# Introduction
## Background
In the context of introductory STEM courses, technical writing training for undergraduates often is poorly designed, implemented, and executed. Often it is treated as a mechanical skill that each student develops "along the way." Also, faculty often are not directly responsible for developing this skill. Technical writing training usually is part of a lab course associated with a lecture class. These lab sections may be be led by graduate teaching assistants or even advanced undergraduates. These class leaders may have under-developed writing skills themselves, or lack pedagogical content knowledge needed to guide students effectively.

@@

*  Focus on holistic coaching 
*  Embody best practices from the writing in the disciplines literature
*  Can be implemented at large scale by less experienced teaching assistants
*  Focus student and instructor attention on larger issues rather than simple copy-correction.

As part of this project, we have assembled a structured database of laboratory reports written by undergraduate students for introductory biology courses. In addition to the students' original texts we have collected these meta-data:

@@

Using text-mining, we are searching for features of student writing, as well as connections and correlations between those features, that can inform how we teach technical writing.

## Context for This Project
It is well established that students make greater gains as writers if instructors adopt a coaching- centric approach and provide holistic, growth-oriented guidance rather than just telling students what to correct. To that end, we train graduate teaching assistants use a **global first, specific later** approach to grading undergraduate technical writing. They are instructed not to spend endless time copy editing. Instead we ask them to spend a limited block of time writing fewer comments that focus on broader issues. Ideally, TAs provide no more than 3-5 comments per text page, and comments should encourage student reflection. As students' writing improves, the TA can focus on smaller issues.

While this approach is sound in theory, we are uncertain how well TAs can implement the recommended strategy. One of our ongoing questions has been, "what do TAs ACTUALLY spend the most time and effort making comments about as they grade student writing?"

>>>>

# Introduction
## Background
Undergraduate introductory STEM courses typically include technical writing training. Often it is treated as an incidental skill that each student develops "along the way." Or, the instructional strategy is  haphazardly designed, implemented, and executed. Adding to the challenge, faculty may not be directly responsible for helping students develop writing skills.

Technical writing training usually is part of a lab course associated with a lecture class. These lab sections may be be led by graduate teaching assistants or even advanced undergraduates. These class leaders may have under-developed writing skills themselves, or lack pedagogical content knowledge needed to guide students effectively.

@@

*  Embody best practices from nearly 40 years of "writing in the disciplines"" literature
*  Can be implemented at large scale by less experienced teaching assistants
*  Emphasize student development through holistic coaching 
*  Focus student and instructor attention on larger issues rather than simple copy-correction.

We have amassed a structured database of laboratory reports written by undergraduate students for introductory biology courses. Besides the original texts we have collected these meta-data:

@@

We are using text-mining approaches to map key features of student writing, as well as connections and correlations between those features, that can inform how we teach technical writing.

## Context for This Project
Students make greater gains as writers when instructors avoid just telling students what to correct ("copy editing"), and adopt a coaching- oriented approach instead. To that end, we train graduate teaching assistants use a **global first, specific later** approach to grading undergraduate technical writing. We ask them to spend a limited block of time writing fewer comments that focus on broader issues. Ideally, TAs provide no more than 3-5 comments per text page, and comments should encourage student reflection. As students' writing improves, the TA can focus on smaller issues.

While this approach is well-grounded in research, we are uncertain how well TAs can implement the recommended strategy. One of our ongoing questions has been, "what do TAs ACTUALLY spend the most time and effort making comments about as they grade student writing?"
